---
title: "Lecture 4 Answers"
format:
  html: 
    theme: literal
---

## Answer 1

```{r}
x <- seq(0,8*pi,length.out=100)
y <- sin(x)
plot(x,y)

```

The x-axis represents the values of the vector x that we created, the
y-axis represents the values of the vector x after the mathematical
function of sine.

------------------------------------------------------------------------

The type indicating the type of plotting. The following changes will
occur accordingly: 

“p” for points.

“l” for lines.

“b” for both points and lines.

“c” for empty points joined by lines.

“o” for overplotted points and lines.

“s” and “S” for stair steps.

“n” does not produce any points or lines.


------------------------------------------------------------------------

using the plot() function, add the pch argument between 1 and 25, In
addition use the rainbow color argument with 5 unique colors.

```{r}
plot(x,y, pch = 1:25, col=rainbow(5))

```

___

```{r, warning=FALSE, message=FALSE}
# Step 1: Call the pdf command to start the plot

pdf(file = "fancy1.pdf")

# Step 2: Create the plot with R code

fancy1 = plot(x,y, pch = 1:25, col=rainbow(5))

# Step 3: Run dev.off() to create the file!

dev.off()

```

## Answer 2

```{r}

boxplot(cars$speed)

```
The line represents the median of speed, which is 15. The box itself shows the upper and lower quartiles. The whiskers show the range (i.e. the largest and smallest values).

___

```{r}
boxplot(speed ~ dist, data = cars,
            xlab = "Stop Distance", ylab = "Speed",
            col = "red")
    title(main="Car Speed Against Stopping Distance", font.main = 3)
```

A summary boxplot with all the guidance.


___

```{r, message=FALSE, warning=FALSE}

library(vioplot)
vioplot(cars$speed)

```


```{r, message=FALSE, warning=FALSE}
library(vioplot)
vioplot(speed ~ dist, data = cars,
            xlab = "Stop Distance", ylab = "Speed",
            col = "red")
    title(main="Car Speed Against Stopping Distance", font.main = 3)
```

## Answer 3

```{r}

whale = read.csv("whaledata.csv")

```

```{r}
par(mar = c(4.0, 4.0, 4.0, 1.9), xaxs="i", yaxs="i")
plot(whale$number.whales, whale$depth,
       type = "n",
       xlab = "Nnumber of whales",
       ylab = expression(paste("Depth (m) ")),
       xlim = c(0, 30), ylim = c(0, 1700), bty = "l",
       las = 1, cex.axis = 0.8, tcl = -0.2)
points(x = whale$number.whales[whale$water.noise == "low"],
       y = whale$depth[whale$water.noise == "low"],
       pch = 16, col = "green")
points(x = whale$number.whales[whale$water.noise == "medium"],
       y = whale$depth[whale$water.noise == "medium"],
       pch = 16, col = "orange")
points(x = whale$number.whales[whale$water.noise == "high"],
       y = whale$depth[whale$water.noise == "high"],
       pch = 16, col = "red")
text(x = 28, y = 1600, label = "A", cex = 2)
leg_cols <- c("green", "orange", "red")
leg_sym <- c(16, 16, 16)
leg_lab <- c("low", "medium", "high")

legend(x = 25, y = 800, col = leg_cols, pch = leg_sym, 
        legend = leg_lab, bty = "n", 
        title = "Noise level")



```



