---
title: "Quiz 1 - 2022/2023"
author: "Introduction to data science for Geographers"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Before we start!

Starting time of the quiz: 9:00

Ending time of the quiz: 11:00

Added time 25% (Only for students with extra time approval) : 11:30

___

**All the code must be submitted in an R file; use # to write the question number and then the answer.**

Example:

```{r,eval=FALSE}
#Question 1

#My code  

x = c(1,2,3,4)

mean(x)


```


**Use # to answer questions that require a verbal answer!!!**

**It is recommended to answer the questions in English. Those who prefer to answer the questions in Hebrew must add a Word file with the answers in Hebrew to the submission box.**

**All files should be saved and submitted with your ID number!
For example:**

`111111111.R`

___

## Part 1 (30 Points)

## Question 1

### How a data frame and a matrix are differing from each other?

A. Unlike data frame, the dimension (dim) of a matrix will always be 2 by 2.

B. Data frame can include different type of data but matrix can contain only similar type of data.

C. Matrix can be consisting only of numbers (numerics), while data frame deals with all data types.

D. It is possible to write a csv file from a data frame, however not from a matrix.


## Question 2

### How many plotting symbols are there as part of the pch argument?

A. 25

B. 15

C. 18

D. 20


## Question 3

### What are Factors in R?

A. Factors are text values on which you can perform mathematical operations.

B. Factors are the data objects which you can perform and categorize ‏binary operations.

C. Factors are the data objects which are used to categorize the data and store it as levels.

D. Factors are a data type designed to handle and order complex numbers.

## Part 2 (70 Points)

The USA cars data set includes information about brands of clean and used vehicles for sale in the US. The data set has the following columns:

* Id - The Id number of the car.

* Price - The sale price of the vehicle in the ad.

* Brand	 - The brand of the car.

* Model	- The model of the vehicle.

* Year - The vehicle registration year.

* Title_status - This feature included binary classification, which are clean   title vehicles and salvage insurance.

* Mileage - The miles traveled by vehicle.

* Color - The color of the vehicle.

* Vin - The vehicle identification number is a collection of 17 characters.   (digits and capital letters)

* State - The location in which the car is being available for purchase.

* Country	- The location in which the car is being available for purchase.

* Sell_time	- Sales target in terms of time.

## Question 1

* Import the USA_cars_datasets.csv file into R.

## Question 2

+ Display the first 10 rows of the data frame.

+ Display the last 10 rows of the data frame.

## Question 3

* Change the name of the "Vin" column to "Vehicle_identification_number".

## Question 4

* Answer the following questions:

A.  What is the total number of NAs in the data set? 
B.  What is the highest Mileage in the data set?  
C.  What is the lowest price in the data set? 
D.  How many "ford" cars are in the data set? 
E.  What is the average car price in Florida?


## Question 5

* From the original USA_cars_datasets.csv, create a new data set named cheap_cars that will be ordered so that the lowest price will be first and the highest price will last.


## Question 6

* From cheap_cars, create a new data frame named Top_50 that will include the first 50 cheapest cars.

* Delete the last column from Top_50.



## Question 7

+ Create a plot of Cars mileage in relation to the Cars price from the **USA_cars_datasets.csv.**

+ The color of the plot should be purple.

+ Add the title of "Cars mileage in relation to the Cars price"

+ Rename the x-axis to "Car Price"

+ Rename the y-axis to "Car mileage"

+ Is there an abnormal dot/dots in the plot?

You should get something like this:


```{r,echo=FALSE}
cars = read.csv("USA_cars_datasets.csv")
plot(cars$Price, cars$Mileage, col = "purple", main="Cars price in relation to car mileage", xlab = "Car Price", ylab = "Car mileage")
```


