---
title: "Lecture 3 Homework Answers"
format:
  html: 
    theme: literal
---


## Question 1

Creating top 10:

```{r}

whaledata = read.csv("whaledata.csv")
TOP10 = whaledata[order(-whaledata$depth),]
TOP10 = TOP10[1:10,]


```

Delete the gradient and noise columns:

```{r}
TOP10 = TOP10[,-c(3,8)]
```

## Question 2

```{r}
#Create a new data frame named noise_location
noise_location = whaledata[,c(3,5:6)]

#Rename water.noise to noise_level
colnames(noise_location)[1] =  "noise_level"

#Order the data frame by noise_level
noise_location$noise_level = factor(noise_location$noise_level,levels = c("low", "medium", "high"))
noise_location = noise_location[order(noise_location$noise_level),]

#Add a column of the average latitude value
noise_location$avg_lat = mean(noise_location$latitude)

#Add a column of the average longitude value
noise_location$avg_long = mean(noise_location$longitude)

write.csv(noise_location, file = "noise_location.csv")
#save the noise_location data frame in csv format to your local folder.


```

## Question 3

```{r, eval=FALSE}
    c(1:10)*10 == seq(10, 100, 10)
    
    rep(10,10) == seq(100,10,-10)
```

In the first code on the left side, all values between 1 and 10 will be multiplied by 10, on the right side, on the other hand, an arithmetic progression series will be created that increases by 10 each time, therefore according to the comparison parameter in the code, all values will receive the value true.

In the second code on the left side the numerical value of 10 will be repeated 10 times, on the right side we will have a descending series between 100 and 10, therefore all the values will receive false except for the last value.

___

```{r}
v <- c(42, 92, 39, 19, 12, 16, 2, 0, 8.5, 50, 4,
           13, 0, -4, 9, NA, 0, 3, 9, 32, 28, 7, 11,
           -3, 8, 74, 70, 12, 20, -1, 6.7, 52, NA, 3, 
           -1, 1, 0, 5, 42, -0.2, 17, 3.1, -9, 11, 1,
           51, 46, 2, 51, 13, NA, 0, 30, -5, 9, 10)

length(v)
# The lengh of the vector

which.max(v)

# Position of the the highest value

which.min(v)

# Position of the the lowest value

```


## Question 4
```{r,eval=FALSE}
#Save your entire workspace, naming it the third_lesson, afterward upload the third_lesson file back into R.

save.image("third_lesson.rdata")

load("third_lesson.rdata")

```

## Bonus question

```{r}

hist(whaledata$depth, breaks = length(whaledata$depth) ,col="blue", main="Histogram of the depth vector") 

```

Although the histogram has a shape with a potential for a normal distribution, the information is not normally distributed, in the upcoming lessons you will learn why.
