---
editor_options: 
  markdown: 
    wrap: 72
---

## Question 1

**Continuation of question 3**

Extract rows from the whale data frame based on the following
criteria:

-   depths greater than 1200 m gradient steeper than 200 degrees water.

-   noise level of 'low' water.noise level of 'high' in the month of
    'May'.

-   The month of 'October', water noise level of 'low' and gradient greater
    than.

-   the median value of gradient (132) all observations from between.

-   latitudes 60.0 and 61.0 and longitudes -6.0 and -4.0.

-   all rows that do not have a water noise level of medium.

-   depths greater than 1500 m and with a greater number of whales
    spotted than average.

## Question 2

As part of our research we want to check what are the ten deepest depths
that whales have reached.

-   Create a new data frame named "TOP10" that contains only the 10
    deepest depths.
-   From "TOP10" delete the gradient and noise columns

## Question 3

As part of our research on whales, we want to create a map with the
location of the whales and their noise level.

-   Create a new data frame named noise_location that contains latitude,
    longitude, and water.noise columns from whale data.

-   Rename water.noise to noise_level

-   Order the data frame by noise_level

-   Add a column of the average latitude value

-    Add a column of the average longitude value

-    save the noise_location data frame in csv format to your local
    folder.

## Question 4

try to guess what you will get before running the following code:

```{r, eval = FALSE}
    c(1:10)*10 == seq(10, 100, 10)
    
    rep(10,10) == seq(100,10,-10)
    
```

## Question 5

Save your entire workspace, naming it the third_lesson, afterward
upload the third_lesson file back into R.

___

## Bonus question! That is for those who want to go further beyond!

Create a histogram of the depth, the histogram should be blue and with an appropriate title, add breaks as the length of the depth vector.

It should look like this:
```{r, echo = FALSE}
whaledata = read.csv("whaledata.csv")
```


```{r, echo = FALSE}

hist(whaledata$depth, breaks = length(whaledata$depth) ,col="blue", main="Histogram of deapth") 

```

