---
title: "Lecture 3 Homework"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

## Question 1

As part of our research we want to check what are the ten deepest depths
that whales have reached.

-   Create a new data frame named "TOP10" that contains only the 10
    deepest depths.
-   From "TOP10" delete the gradient and noise columns

## Question 2

As part of our research on whales, we want to create a map with the
location of the whales and their noise level.

-   Create a new data frame named noise_location that contains latitude,
    longitude, and water.noise columns from whale data.

-   Rename water.noise to noise_level

-   Order the data frame by noise_level

-   Add a column of the average latitude value

-   Add a column of the average longitude value

-   save the noise_location data frame in csv format to your local
    folder.

## Question 3

try to guess what you will get before running the following code:

```{r, eval = FALSE}
    c(1:10)*10 == seq(10, 100, 10)
    
    rep(10,10) == seq(100,10,-10)
    
```

------------------------------------------------------------------------

Create the following vector:

```{r}
v <- c(42, 92, 39, 19, 12, 16, 2, 0, 8.5, 50, 4,
           13, 0, -4, 9, NA, 0, 3, 9, 32, 28, 7, 11,
           -3, 8, 74, 70, 12, 20, -1, 6.7, 52, NA, 3, 
           -1, 1, 0, 5, 42, -0.2, 17, 3.1, -5, 11, 1,
           51, 46, 2, 51, 13, NA, 0, 30, -5, 9, 10)

```

-   What is the length of the vector?

-   How many many Na values are there in the vector?

-   In which position in the vector is the the highest value?

-   In which position in the vector is the the lowest value?

## Question 4

Save your entire workspace, naming it the third_lesson, afterward upload
the third_lesson file back into R.

## Bonus question! That is for those who want to go further beyond!

Create a histogram of the depth, the histogram should be blue and with
an appropriate title, add breaks as the length of the depth vector.

It should look like this:

Does the histogram have a normal distribution?

```{r, echo = FALSE}
whaledata = read.csv("whaledata.csv")
```

```{r, echo = FALSE}

hist(whaledata$depth, breaks = length(whaledata$depth) ,col="blue", main="Histogram of the depth vector") 

```
