---
title: "Lecture 2 Exercises"
format:
   html:
    self-contained: true
    theme: literal
---

## Before we start!

Paste the lesson files into your home directory. Run the following code
in Git Bash:

```{r,eval=FALSE}
cd ~/ADS_L2

```

## Question 1

The CLI has survived and thrived for almost fifty years because it lets
people do complex things with just a few keystrokes. It helps users
combine existing programs in new ways, automate repetitive tasks, and
run programs on clusters and clouds that may be halfway around the
world. This lesson will introduce its key elements and show you how to
use them efficiently.

As part of your internship in the world of data, you were privileged to
manage a company's database. You want and need to manage the data in the
best possible way.

### 1.1 The basis of any successful code is to know which files we are working with and navigate through directories.

1.  Run `pwd` command (Print working directory), where are you right
    now?

2.  Run the `ls -la` command what do you get?

3.  Change the directory to the Report folder, what files are there?

4.  Where does cd . takes you?

5.  Where does cd .. takes you?

6.  Where does cd \~ takes you?

7.  Return back to the Report folder, where does cd \~/../. take you?

## 1.1 Answer

Run the pwd command (Print working directory), where are you right now?

You should be in \~/ADS_L2 inside the lesson directory you copied to
your home directory.

___

Run the ls -la command what do you get?

ls is a Linux shell command that lists directory contents of files and
directories. In addition the the -l and -a argummnets will show you the
followong:

-l: Will show a "long listing", which includes extra details such as
permissions, ownership, file sizes, and timestamps

-a: WIll list all of a directory's files, including hidden ones (that
start with .)

___

Change the directory to the Report folder, what files are there?

Q1.csv 

Q2.csv

Q3.csv 

Q4.csv

___

where does cd . take you?

This command will keep you in your current directory.

where does cd .. take you?

This command will take you one directory back.

where does cd \~ take you?

This command will take you to the home directory back.

___


Return back to the Report folder, where does cd \~/../. take you?

This command will return you one directory before the home directory.

___

### 1.2 Now we will focus on creating new data and deleting unnecessary data.

1.  Create a new folder named backup, Make a copy of Q1.csv in the
    backup folder, and call the new file Q1.bck.

2.  Copy Q2.csv and Q3.csv from the Report directory into the backup
    directory without changing your current working directory.

3.  Using a single command, move Q1.csv and Q4.csv from Report to the
    backup folder.

4.  Rename Q4.csv to Autumn.csv

5.  Remove Autumn.csv.

6.  Without changing directories, delete the file info.txt in the
    More_info directory.

7.  Now that the More_info directory is empty, use a single command to
    delete it.

8.  Create a new directory called yearly below your home directory. Now that annually exists, create another directory called 2017 inside of it without leaving your home directory.
    

## Answer 1.2

```{r,eval=FALSE}

Create a new folder named backup, Make a copy of Q1.csv in the backup folder, and call the new file Q1.bck.

cd ~/ADS_L2

mkdir Backup

cp  ~/ADS_L2/Report/Q1.csv ~/ADS_L2/Backup/Q1.bck

```

```{r,eval=FALSE}

Copy Q2.csv and Q3.csv from the Report directory into the backup
directory without changing your current working directory.

cp  ~/ADS_L2/Report/Q2.csv ~/ADS_L2/Report/Q3.csv Backup

```

```{r,eval=FALSE}
Using a single command, move Q1.csv and Q4.csv from Report the backup folder.

mv  ~/ADS_L2/Report/Q1.csv ~/ADS_L2/Report/Q4.csv Backup



```

```{r,eval=FALSE}

Rename Q4.csv to Autumn.csv

cd ~/ADS_L2/Backup

mv Q4.csv Autumn.csv

```

```{r,eval=FALSE}

Remove Autumn.csv.

rm Autumn.csv

```

```{r,eval=FALSE}

Without changing directories, delete the file info.txt in the More_info directory.

rm info.txt ~/ADS_L2/More_info/info.txt

```

```{r,eval=FALSE}

Now that the More_info directory is empty, use a single command to delete it.

rmdir ~/ADS_L2/More_info

```

```{r, eval=FALSE}
Create a new directory called yearly below your home directory. Now that annually exists, create another directory called 2017 inside of it without leaving your home directory.

Option A :
  
cd ~/ADS_L2
  
mkdir yearly

mkdir ~/ADS_L2/yearly/2017
  
Option B :
  
mkdir -p yearly/2017

```

### 1.3 Now we will focus on displaying data

1.  View the content of Q3.csv from the Backup directory.

2.  Display only the first 3 lines of Q3.csv.

3.  Read the manual page for the tail command to find out what putting
    a + sign in front of the number used with the -n flag does.
    (Remember to press the spacebar to page down and/or type q to quit.)

4.  Use tail with the flag -n +7 to display, what do you get?

## Answer 1.3

```{r,eval=FALSE}

View the content of Q3.csv from the Backup directory. 

cd ~/ADS_L2/Backup

cat Q3.csv

```

```{r,eval=FALSE}

Display the first 3 lines of Q3.csv.

head -n 3 Q3.csv
```

```{r,eval=FALSE}

Use tail with the argument of -n +7, what do you get?
  
  
tail -n +7 Q3.csv

The command "tail -n +7 Q3.csv" will display the contents of the file "Q3.csv" starting from the 7th line till the end of the file.

```

## Question 2

Head and tail let you select rows from a text file. If you want to
select columns, you can use the command cut. It has several options (use
man cut to explore them), but the most common is something like:

cut -f 3-6,9 -d , values.csv which means "select columns 3 through 6 and
columns 9, using comma as the separator". cut uses -f (meaning "fields")
to specify columns and -d (meaning "delimiter") to specify the
separator. You need to specify the latter because some files may use
spaces, tabs, or colons to separate columns.

#### Question 2.1 What command will select the first column (containing dates) from the file Q3.csv?

## Answer 2.1

```{r, eval=FALSE}
What command will select the first column (containing dates) from the file Q3.csv?

cd ~/ADS_L2/Backup
  
cut -d , -f 1 Q3.csv

```

------------------------------------------------------------------------

How can I select lines containing specific values? head and tail select
rows, cut selects columns, and grep selects lines according to what they
contain. In its simplest form, **grep** takes a piece of text followed
by one or more filenames and prints all of the lines in those files that
contain that text.

Here are some common arguments:

-   c: print a count of matching lines rather than the lines themselves.

-   h: do not print the names of files when searching multiple files.

-   i: ignore case (e.g., treat "Regression" and "regression" as
    matches).

-   l: print the names of files that contain matches, not the matches.

-   n: print line numbers for matching lines.

-   v: invert the match, i.e., only show lines that don't match.

### Question 2.2 Follow the instructions below:

1.  Print the contents of all of the lines containing the word Italy in
    Forbes2000.csv

2.  Invert the match to find all of the lines that don't contain the
    value 2013 in Q3.csv.

3.  Count how many lines contain the the word Italy in Forbes2000.csv.

4.  Combine tail with redirection to save the last 3 lines of Q3.csv in
    a file called last_3.csv.

5.  With one command, copy all the text files to the backup directory.

## Answer 2.2

```{r,eval=FALSE}


Print the contents of all of the lines containing the value Italy in Forbes2000.csv


cd ~/ADS_L2/More_info2/

grep Italy Forbes2000.csv
```

```{r,eval=FALSE}

Invert the match to find all of the lines that don’t contain the value 2013 in Q3.csv.

cd ~/ADS_L2/Backup/

grep -v -n 2013 Q3.csv
```

```{r,eval=FALSE}

Count how many lines contain the the word Italy in Forbes2000.csv

cd ~/ADS_L2/More_info2/

grep -c Italy Forbes2000.csv

```

```{r,eval=FALSE}

Combine tail with redirection to save the last 3 lines of Q3.csv in a file called last_3.csv.

cd ~/ADS_L2/Backup/

tail -n 4 Q3.csv > last_3.csv  

```

```{r,eval=FALSE}
With one command, copy all the text files from More_info2 to the backup directory.

cp ~/ADS_L2/More_info2/*.txt ~/ADS_L2/Backup

```

## Question 3

Unix has a variety of text editors. For this course, we will use a
simple one called Nano. You can move around with the arrow keys, delete
characters using backspace, and do other operations with control-key
combinations:

Ctrl + K: delete a line.

Ctrl + U: un-delete a line.

Ctrl + O: save the file ('O' stands for 'output'). You will also need to
press Enter to confirm the filename!

Ctrl + X: exit the editor.

### 3. Follow the instructions below:

1.  Add your name and your birthday in the words.txt file.

## Answer 3

```{r,eval=FALSE}

cd ~/ADS_L2/More_info2

nano words.txt
```

## Lecture 2 Homework

## Question 1

Inspect the following code:

```{r, eval=FALSE}
cd ~
touch my_file.txt  
```

### 1. Follow the instructions below:

1.  What did the touch command do?

2.  When you look at your home directory using the GUI file explorer,
    does the file show up?

3.  Use ls -l to inspect the files. How large is my_file.txt?

4.  When might you want to create a file this way?

## Answer 1

##### What did the touch command do?

The touch command generates a new file called 'my_file.txt' in your home
directory.

##### When you look at your home directory using the GUI file explorer, does the file show up?

If you are in your home directory, you can observe this newly generated
file by typing 'ls' at the command line prompt.

'my_file.txt' can also be viewed in your GUI file explorer.

##### Use ls -l to inspect the files. How large is my_file.txt?

When you inspect the file with 'ls -l', note that the size of
'my_file.txt' is 0kb.

In other words, it contains no data. If you open 'my_file.txt' using
your text editor it is blank.

##### When might you want to create a file this way?

Some programs do not generate output files by themselves but instead
require empty files that have already been generated.

When the program is running, it searches for an existing blank file to
pipe the output to it.

The touch command allows you to efficiently generate a blank text file
to be used by such programs.

## Question 2

What happens when we type rm -i \~/ADS_L2/More_info2/words.txt

Why would we want this protection when using rm?

## Answer 2

The -i argument will display a prompt before every removal.

The Unix shell doesn't have a trash bin, so all the files removed will
disappear forever.

By using the -i flag, we have the chance to check that we are deleting
only the files that we want to remove.

## Question 3

What is the output of the code shown below?

```{r,eval=FALSE}
$ pwd
/Users/jamie/data
$ ls
proteins.dat
$ mkdir recombine
$ mv proteins.dat recombine/
$ cp recombine/proteins.dat ../proteins-saved.dat
$ ls
```

Is it:

1.  proteins-saved.dat recombine

2.  recombine

3.  proteins.dat recombine

4.  proteins-saved.dat

## Answer 3

We start in the /Users/jamie/data directory, and create a new folder
called recombine.

The second line moves (mv) the file proteins.dat to the new folder
(recombine).

The third line makes a copy of the file we just moved.

The tricky part here is where the file was copied to.

Recall that .. means "go up a level", so the copied file is now in
/Users/jamie.

Notice that .. is interpreted with respect to the current working
directory, not with respect to the location of the file being copied.

So, the only thing that will show using ls (in /Users/jamie/data) is the
recombine folder - Option 2.

## Question 4

Jamie is working on a project and she sees that her files aren't very
well organized:

```{r,eval=FALSE}
$ ls -F
#analyzed/  fructose.dat    raw/   sucrose.dat
```

The fructose.dat and sucrose.dat files contain output from her data
analysis.

What command(s) does she need to run so that the commands below will
produce the output shown?

```{r,eval=FALSE}
$ ls -F
analyzed/   raw/

$ ls analyzed
fructose.dat    sucrose.dat


```

## Answer 4

```{r,eval=FALSE}
mv *.dat analyzed

```

Jamie needs to move her files fructose.dat and sucrose.dat to the
analyzed directory.

The shell will expand \*.dat to match all .dat files in the current
directory.

The mv command then moves the list of .dat files to the "analyzed"
directory.

## Bonus question! That is for those who want to go further beyond!

##### Another essential part of the shell is writing code with variables, solve the following questions.

Let us have a look at the syntax of the if-else code.

```{r,eval=FALSE}
if [condition]
then
   statement1
else
   statement2
fi

```

1.  Using if-else write a code to check whether a number is even or not.

2.  Using if-else write a code to check whether two numbers are equal or
    not. (hint use -eq in the if condition)

Here is a nice site for writing the code:

<https://www.onlinegdb.com/online_bash_shell>

## Bonus question Answer

1.  

```{r,eval=FALSE}
echo "---- EVEN OR ODD IN SHELL SCRIPT -----"
echo -n "Enter a number:"
read n
echo -n "RESULT: "
if [ `expr $n % 2` == 0 ]
then
	echo "$n is even"
else
	echo "$n is Odd"
fi
```

or:

```{r,eval=FALSE}
n=9
if [ $((n%2)) -eq 0 ]
then
  echo "The number is even."
else
  echo "The number is odd."
fi
```

2.  

```{r,eval=FALSE}

q=1
w=2

if [ $q -eq $w ]
then
        echo "Both variables are the same"
else
        echo "Both variables are different"
fi

```
