---
title: "Lesson 5 Homework"
format:
  html: 
    theme: literal
---

## Question 1

For the next research exercise, we will use the data set about mice.

-   Download and read the mice.csv data set and take a first look at the data.

-   How large is the data.frame? use only 1 function to get the answer.

-   how many rows/columns?

-   what are the column names and what is the data type of each column?

-   How many mice are described in the data set?

-   How many female mice are there?

-   Make a plot of the weight of mouse number 1

-   Make a histogram of the weight of mouse number 5

## Question 2

Copy and run the following vector:

```{r}
v <- c(42, 92, 39, 19, 12, 16, 2, 0, 8.5, 50, 4,
           13, 0, -4, 9, 10, 0, 3, NA, 32, 28, 7, 11,
           -3, 8, 74, 70, 12, 20, -1, 6.7, NA, 6, 3, 
           -1, 1, 0, NA, 42, -0.2, 17, 3.1, -5, 11, 1,
           51, 46, 2, 51, 13, NA, 0, 30, -5, 9, NA)
```

Print five-number summary using one function, and afterward analyze the result.

## Bonus question! That is for those who want to go further beyond!

-   From the built-in data set VADeaths, create a bar plot with a legend of the age groups.

you should get:

```{r, echo = FALSE}
barplot(VADeaths, legend= rownames(VADeaths))

```

-   In addition, unstack the bars and plot them alongside one another, adding to each age group color and a proper title.

you should get:

```{r, echo = FALSE}
barplot(VADeaths, beside = TRUE, 
    col = c("lightblue", "mistyrose", "lightcyan","lavender", "cornsilk"),
    legend = rownames(VADeaths), ylim = c(0, 100))
    title(main = "Death Rates in Virginia", font.main = 4)
```
