---
title: "Lecture 5 Exercises"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: sentence
---

## Question 1

Paste the following code in R:

```{r}
x <- c(7,5,2,11,2,3,6,0,1,4,8,9,5,12,11,5,1,0,2,2,7,3,10,19,18,
          9,20,14,15,1,16,9,13,0,0,2,11,3,4,1,7,9,10,5,3,4,13,15,
          19,14,2,4,1,12,15,12,11,10,7,4,11,8,1,11,2,3,0,12,17,21)
y <- c(1:3)
```

-   How many elements are in x?

-   Try to think what will be the result of: y \> 2 & y-1 == 2

-   How many elements in x are equal or bigger than 10?

-   How many elements in x are bigger than 8 and smaller than 14?

------------------------------------------------------------------------

```{r}
type <- c("Smart", "3D", "Smart", "Smart", "4K", "4K", "3D", "Smart", "4K", "3D")
    size <- c(45, 50, 45, 40, 50, 50, 40, 38, 50, 50)
    price <- c(3000, 4200, 3500, 3000, 5200, 5100, 3800, 2400, 5100, 4200)
```

-   Create a data frame from the vectors above.

-   Use `Duplicated()` and `Unique()` to see how many repeated observations are in the data

-   Create a cleaned data frame without the repeated observations

## Question 2

The Forbes2000 list is a ranking of the world's biggest companies in 2004, measured by sales, profits, assets and market value.

The data format is a data frame with 2000 observations on the following 8 variables:

1.  rank - the ranking of the company.

2.  name - the name of the company.

3.  country - a factor giving the country the company is situated in.

4.  category - a factor describing the products the company produces.

5.  sales - the number of sales of the company in billion USD.

6.  profits - the profit of the company in billion USD.

7.  assets - the assets of the company in billion USD.

8.  marketvalue - the market value of the company in billion USD.

------------------------------------------------------------------------

-   Download and read the forbes2000 file in R

-   What is the maximum value of assets?

-   What is the minimum value of sales?

-   What is the mean of profits?

-   How many NA's there are in profits?

-   Is there any company that appears twice in the forbes list?

## Question 3

Use the summary tool and see how can you answer the previous questions without any trouble.

-   Use summary only on companies with higher market value than the market value mean (of all companies)

-   Use rescale to create a new column named marketvalue.rescaled in the forbes2000 data. (use the package scales)

-   Use `cut()` to create the column marketvalue_cat, so you will have 3 categorical bins of "Low", "Medium" and "High" for the marketvalue

-   Create and combine histograms for profits, assets and sales and see the differences
