---
title: "Lecture 9 Exercises"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

It is often said that 80% of data analysis is spent on cleaning and
preparing data. Furthermore, it is not only the first step in each data
science project, but a process that must be repeated throughout the
analysis as new problems come to light or new data is collected. To get
a handle on the problem, this exercise focuses on an essential aspect of
data called data tidying: structuring datasets to facilitate analysis.

-   The general principle of "tidy data" is this:

-   Every column is a variable Every row is an observation

-   Every cell is a single value.

## Question 1

### 1.1 - relig_income data tibble

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
relig_income
```

relig_income is a tibble that contains the result of the religion and
income survey.

1.  Why is the data in the tibble not tidy?

2.  Tidy the relig_income tibble.

### 1.2 - fish_encounters tibble

```{r}
fish_encounters
```

fish_encounters is a tibble containing information about fish swimming
down a river. Each station represents an autonomous monitor that records
if a tagged fish was seen at that location.

1.  Why is the data in the tibble not tidy?

2.  Tidy the fish_encounters tibble.

3.  What problem appeared in the new tibble? Try to fix it.

## Answer 1.1

The issue is that all the columns except for religion aren't really
variables. They are realizations of a different variable - that is,
income. So we need to make this tibble go from "wide" (more columns) to
"long" (fewer columns, more rows). Our function for this in tidyr is
`pivot_longer()`.

```{r}
relig_income %>%
  pivot_longer(cols = -religion, names_to = "income", values_to = "value")
```

## Answer 1.2

fish_encounters has the opposite problem of relig_income. The variables
"Release" and "I80_1," etc., are condensed under one variable,
"station." In essence, not every row is an observation that holds other
variables fixed. We need to go from having a "long" dataset with many
rows to a "wider" dataset that has more columns. We will use the
`pivot_wider()` function for this.

```{r}
fish_encounters %>%
  pivot_wider(names_from = station, values_from = seen)
```

------------------------------------------------------------------------

In the new tibble, we have plenty of NA values, and we can fix it with a
small adjustment to the previous code.

```{r}
fish_encounters %>%
  pivot_wider(names_from = station, values_from = seen, values_fill = 0)
```

## Question 2

```{r}
df <- data.frame(Player=c('Frodo Baggins', 'Frodo Baggins', 'Samwise Gamgee', 'Samwise Gamgee', 'Grover Cleveland', 'Grover Cleveland'),
                 Year=c(1999, 2000, 1999, 2000, 1999, 2000),
                 Points=c(31, 22, 30, 15, 16, 17),
                 Assists=c(3, 5, 4, 6, 2, 1))
```

Sometimes we want to turn two variables into one single variable or one
variable into several variables.

1.  Combine the "points" and "assists" columns into a single column
    called Player_Stat.

2.  Separate the Player column to Personal Name and Family Name.

## Answer 2

Combine the "points" and "assists" columns into a single column called
Player_Stat.

```{r}
df %>% 
unite(col='Player_Stat', c('Points', 'Assists'), sep='-')

```

------------------------------------------------------------------------

Separate the Player column to Personal Name and Family Name.

```{r}
df %>% 
unite(col='Player_Stat', c('Points', 'Assists'), sep='-') %>%
separate(Player, into = c("Personal Name", "Family Name"), sep = " ", convert = TRUE)  

```

## Question 3

```{r}
tidy_pets <- tibble::tribble(
  ~rank, ~pet_type, ~breed,
  1L,        NA,    "Boston Terrier",
  2L,        NA,    "Retrievers (Labrador)",
  3L,        NA,    "Retrievers (Golden)",
  4L,        NA,    "French Bulldogs",
  5L,        NA,    "Bulldogs",
  6L,     "Dog",    "Beagles",
  1L,        NA,    "Persian",
  2L,        NA,    "Maine Coon",
  3L,        NA,    "Ragdoll",
  4L,        NA,    "Exotic",
  5L,        NA,    "Siamese",
  6L,     "Cat",    "American Short"
)

tidy_pets
```

1.  What values are we missing in tibble?

2.  Complete the missing values.

## Answer 3

We can see that there are many missing values regarding the pet type.

```{r}
tidy_pets %>%
  fill(pet_type, .direction = "up")

```

## Question 4

```{r}
head(iris)
```

This famous (Fisher's or Anderson's) iris data set gives the
measurements in centimeters of the variables sepal length and width and
petal length and width, respectively, for 50 flowers from each of the
three species of iris. The species are Iris setosa, versicolor, and
virginica.

1.  Convert the data frame into a tibble.

2.  Tidy the data - create one column of Measures.

3.  Separate the data in Measures into "Part" and "Measure."

4.  Using `geom_jitter()`, plot the Species in relation to the Values.
    The color needs to be the "Part." Facet by "Measure."

You should get something like that:

```{r, echo=FALSE}
iris %>%
  pivot_longer(cols = -Species, names_to = "Measures", values_to = "Values") %>%
  separate(col = Measures, into = c("Part", "Measure")) %>%
  ggplot(aes(Species, Values, color = Part)) + 
    geom_jitter() + 
    facet_grid(~Measure)
```

## Answer 4

Convert the data frame into a tibble.

```{r}
iris2 = as_tibble(iris)

```

------------------------------------------------------------------------

Tidy the data - create one column of Measures.

```{r}
iris2 %>%
  pivot_longer(cols = -Species, names_to = "Measures", values_to = "Values")
```

------------------------------------------------------------------------

Separate the data in Measures into "Part" and "Measure."

```{r}
iris2 %>%
  pivot_longer(cols = -Species, names_to = "Measures", values_to = "Values") %>%
  separate(col = Measures, into = c("Part", "Measure"))
```

------------------------------------------------------------------------

Using `geom_jitter()`, plot the Species in relation to the Values. The
color needs to be the "Part." Facet by "Measure."

```{r}
iris2 %>%
  pivot_longer(cols = -Species, names_to = "Measures", values_to = "Values") %>%
  separate(col = Measures, into = c("Part", "Measure")) %>%
  ggplot(aes(Species, Values, color = Part)) + 
    geom_jitter() + 
    facet_grid(~Measure)
```

## Lecture 9 Homework

## Question 1

We are working on table1, table2, table3, table4a from the lecture.

Are those tables tidy, or aren't they tidy?

## Answer 1

```{r}
table1
```

Table 1 is tidy. All columns have their unique variable; for example,
the year column contains only years, and the country column contains
only countries.

For example, if we got a data set where we have country and then two
columns for males and females, that would be untidy because both columns
should be in a column called gender.

------------------------------------------------------------------------

```{r}
table2
```

Table2 is not a tidy dataset. If we had something like different
sicknesses in the type column, this would be tidy, but the type, in this
case, is not tidy. So we have two different variables inside the type
column.

------------------------------------------------------------------------

```{r}
table3
```

Table3 is non-tidy because we cannot work with values in the rate
column; the values are characters and not numeric.

\_\_

```{r}
table4a
```

Table4a is not a tidy dataset because of the "1999" and "2000" columns.
Both of these columns should be under a "Year" column.

## Question 2

Calculate the rate of illnesses in table2.

You will need to perform six steps:

1.  Extract the cases values for each country by year (use filter).

2.  Extract the population value for each country by year.

3.  Combine the two tables into one table (use bind_cols).

4.  Create the rate column by dividing cases by population and
    multiplying by 10000.

5.  From the tables you created, Delete the unneeded columns.

6.  Bind the two tables into one using bind_rows.

## Answer 2 

### 1 

```{r}
library(tidyverse)
fyear_cases_1999 =
  table2 %>%
  filter(year == 1999,
         type == "cases")


fyear_cases_2000 =
  table2 %>%
  filter(year == 2000,
         type == "cases")

```

___

### 2

```{r}
fyear_pop_1999 =
  table2 %>%
  filter(year == 1999,
         type == "population")

fyear_pop_2000 =
  table2 %>%
  filter(year == 2000,
         type == "population")

```

___

### 3 

```{r}
rate_1999 =
  bind_cols(fyear_cases_1999, fyear_pop_1999)
rate_2000 =
  bind_cols(fyear_cases_2000, fyear_pop_2000)
```

___

### 4

```{r}
rate_1999 <- rate_1999 %>%
  mutate(rate = (count...4/count...8)*10000)


rate_2000 <- rate_2000 %>%
  mutate(rate = (count...4/count...8)*10000)

```

___

### 5 

```{r}
rate_1999 <- rate_1999 %>%
  select(country...1,year...2,count...4,rate )

rate_2000 <- rate_2000 %>%
  select(country...1,year...2,count...4,rate )

```

___

### 6 

```{r}
bind_rows(rate_1999, rate_2000)

```

### Bonus Question  

Create the following data frame:

```{r}
dfz <- tibble(
  x = 1:3,
  y = c("a", "d,e,f", "g,h"),
  z = c("1", "2,3,4", "5,6")
)

dfz
```

___

Separate the rows to get the following result:

```{r,echo=FALSE}
separate_rows(dfz, y, z, convert = TRUE)

```

## Bonus Answer

```{r}
separate_rows(dfz, y, z, convert = TRUE)

```

