---
title: "Lecture 11 Exercises"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

## Question 1

In this exercise, you will combine everything that you've learned so
far: Do you remember Daniel from the first class with the GIS and
ecology companies? Every company has a website that he manages (a very
hardworking guy). Let's help him write code that will give him a summary
of views on each web page.

Here are pre-defined two vectors: GIS and Ecology, They are the number
of views on each site on the last day.

```{r}
GIS = 30
Ecology = 19
```

-   If both GIS and Ecology are 30 or higher, set Views equal to double
    the sum of GIS and Ecology.

-   If both GIS and Ecology are strictly below 20, set Views equal to
    half the sum of GIS and Ecology.

-   In all other cases, set Views equal to GIS and Ecology.

-   Finally, print the resulting Views variable.

A small reminder of what the syntax looks like:

if (condition1) {

expr1

} else if (condition2) {

expr2

} else if (condition3) {

expr3

} else {

expr4

}

## Question 2

A for loop is used to iterate over a vector in R programming.

```{r}
q = 1
```

-   Create a new loop that multiplies the value of q by 4 each time, as
    long as q is less than 16, after each multiplication the next q
    should increase by 1.

-   The loop is stopped with a brake argument if the multiplication by 4
    of q is divisible by 8, but still, print out the result before the
    breaking.

A small reminder of what the syntax looks like:

while (condition) {

  expr
  
}

## Question 3

Who doesn't like to play tic tac toe? Here is a matrix of a tic tac toe game between students during an important lecture.

```{r}
x = c("O","NA","X")
y = c("NA","X","O")
z = c("X","O","O")

m = matrix(c(x, y, z), ncol = 3)
```

By using a pair of for loops print the values in the tic tac toe matrix.
Inside the inner loop, make use of `print(`) and `paste()` to print out information in the following format: "On row x and column y the board contains z", where x is the value on that position.


A small reminder of what the syntax looks like:

for (var1 in seq1) {

  for (var2 in seq2) {
  
    expr
    
  }
  
}

You should get something like that:

[1] "On row 1 and column 1 the board contains O"

[1] "On row 1 and column 2 the board contains NA"

[1] "On row 1 and column 3 the board contains X"

[1] "On row 2 and column 1 the board contains NA"

[1] "On row 2 and column 2 the board contains X"

[1] "On row 2 and column 3 the board contains O"

[1] "On row 3 and column 1 the board contains X"

[1] "On row 3 and column 2 the board contains O"

[1] "On row 3 and column 3 the board contains O"


## Question 4

Create a function to calculate the area of a circle. Test the function by finding the area of a circle with a diameter of 3.4 cm. Can you use it on a vector of data? In addition, make the default diameter of the function to be 1 cm.






