---
title: "Lecture 2 Homework"
format:
  html: 
    theme: literal
---

## Question 1

```{r}
#| echo: true
#| eval: false
random_vector = sample(1:10, size = 6)

random_vector

#This code creates a vector with random values between 1 and 10, feel free to run the code as many times as you want to get new numbers.
```

Run the following code, from random_vector and use a logical index to find the following values:

-   values less than 8

-   values less than or equal to 7

-   values equal to 5

-   values not equal to 6

-   extract values in m which are less than 6 AND greater than 2

-   extract values in m which are less than 9 AND greater than 5

-   replace the 5th and 6th values with 1

-   replace values that are less than or equal to 4 with 5

-   use `sort()` and `order()` functions on random_vector , what is the difference between the two functions?

## Question 2

As part of Sarah's research on US Presidents, she created two separate data frames, help her connect the two data frames using only **one function**.

```{r}
usa_presidents = data.frame(year=c(1789,1829,1861,1980,2000),
order = c(1,7,16,40,43))
names = data.frame(president_name =
c("George_Washington","Andrew_Jackson","Abraham_Lincoln","Ronald Reagan","George_W._Bush"))

```

## Question 3

As part of a population survey you received two data frames, combine the two data frames into one data frame using **only one function**.


```{r}
Student_Name<-c("John","Barak","Sara","Yuval","Muhammad")
Student_class<-c("High school","College","High school","High school","College")
df1<-data.frame(Student_class,Student_Name)


Student_Name<-c("Tal","Amir","Noya")
Student_class<-c("High school","College","University")
df2<-data.frame(Student_Name,Student_class)

```


## Bonus question! That is for those who want to go further beyond!

Try to think by yourself first what the result of the following code will be, then run the code to check if you were right.

```{r, eval = FALSE}
c(1, 2, 3) + c(1, 2, 3, 4, 5, 6)
```

```{r, eval = FALSE}
counts = c(2, 0, 3, 1, 3, 2, 9, 0, 2, 1, 11, 2)

counts[c(1:3, 7:9)] = c(NA, 99)
```

```{r, eval = FALSE}
paste("I", "want", "to", "combine", "text", "values")


```
