---
title: "Lesson 2 Homework Answers"
format:
  html: 
    theme: literal
---

## Question 1 Answer

```{r}
random_vector = sample(1:10, size = 6)

random_vector
```

```{r}
random_vector[random_vector>8]
random_vector[random_vector<=7]
random_vector[random_vector == 5]
random_vector[random_vector!=6]
random_vector[random_vector<6 & random_vector >2]
random_vector[random_vector<9 & random_vector >5]
random_vector[random_vector<=4] = 5
random_vector
```

___

```{r}

sort(random_vector)

order(random_vector)
```

`sort()` sorts the vector in an ascending order, however `order()` returns the indices of the vector in a sorted order.

## Question 3 Answer

```{r}
usa_presidents = data.frame(year=c(1789,1829,1861,1980,2000),
order = c(1,7,16,40,43))
names = data.frame(president_name =
c("George_Washington","Andrew_Jackson","Abraham_Lincoln","Ronald Reagan","George_W._Bush"))
```

```{r}
cbind(usa_presidents,names)
```

## Question 4 Answer

```{r}
Student_Name<-c("John","Barak","Sara","Yuval","Muhammad")
Student_class<-c("High school","College","High school","High school","College")
df1<-data.frame(Student_class,Student_Name)


Student_Name<-c("Tal","Amir","Noya")
Student_class<-c("High school","College","University")
df2<-data.frame(Student_Name,Student_class)
```

```{r}
rbind(df1,df2)
```

## Bonus Question Answer

```{r}
c(1, 2, 3) + c(1, 2, 3, 4, 5, 6)
```
When one of the vectors is of length 3 and the other is of length 6, the shorter vector (of length 3) is replicated 2 times.

```{r}
counts = c(2, 0, 3, 1, 3, 2, 9, 0, 2, 1, 11, 2)

counts[c(1:3, 7:9)] = c(NA, 99)
```

Here, c(NA,99) is replicated three times, also to match the subset of length 6.

```{r}
paste("I", "want", "to", "combine", "text", "values")
```

The proper way to combine text values in R. 



