---
title: "Lecture 8 Exercises"
format:
  html:
    self-contained: true
editor_options: 
  markdown: 
    wrap: 72
---

This week's research will focus on analyzing information about football
players from England, Italy, and Germany. Read football_info.csv into R.
The data consist of information on 15,000 football players.

football_info.csv has the following variables: age, height, weight and
foot (which is the strongest foot of the player), Country, and the
player score from 0 to 100.

## Question 1

### 1. follow the instructions below.

1.  Read football_info.csv into R.

2.  As you know by now, the first thing we do with a new dataframe is to
    examine it. Check the different values and variables in the table,
    look for NA and delete outliers if you notice them.

3.  By using a histogram and QQ-plot, which of the continuous variables
    in the data is normally distributed? (Hint! you can use the
    `qqnorm()` and the `qqline()` functions).

## Answer 1

```{r}
#Read football_info.csv into R

info = read.csv("football_info.csv")

```

------------------------------------------------------------------------

```{r}
#Check and clean the data if required
summary(info)
#With summary(info) we can see that there are no NAs, however, there are some wrong values in a few columns.
#The maximum age is 244, which makes no sense, so we need to delete it.
#We can also notice the impossible maximum value of height (281), the maximum value of weight(909), and the minimum value of weight(3).
#So there are some errors in the data and we have to delete the wrong rows.
info = info[order(info$Age),]
info = info[-c(14999,15000),]
info = info[order(info$Height),]
info = info[-14998,]
info = info[order(info$Weight),]
info = info[-14997,]
info = info[-1,]
summary(info)
#Now our data is fully cleaned :)


```

------------------------------------------------------------------------

```{r}
#Which of the continuous variables in the data is normaly distributed?

#Height

hist(info$Height, xlab = "Age", col="burlywood4", main = "Football Players Height Distribution")
qqnorm(info$Height)
qqline(info$Height, col="burlywood4")

```

------------------------------------------------------------------------

```{r}
#Which of the continuous variables in the data is normaly distributed?

#Weight

hist(info$Weight, xlab = "Age", col="Blue", main = "Football Players Weight Distribution")
qqnorm(info$Weight)
qqline(info$Weight, col="red")
```

------------------------------------------------------------------------

```{r}
hist(info$Age, col = "azure3", xlab = "Age", main = "Football Players Age Distribution")
abline(v = mean(info$Age), col = "blue4")  # With "v=mean(info$Height)" you can add the mean line to the plot.
qqnorm(info$Age)
qqline(info$Age, col="red")
```

You cannot be sure that the data is normally distributed with plots.
However, you can rule out if it is NOT normally distributed. All three
variables have a good chance of being normally distributed.

## Question 2

### 2. follow the instructions below.

The number of standard deviations from the mean is called the "Standard
Score" or "Z-Score". "Standardizing" is converting a value to a Standard
Score.

1.  Create the column "Standard_Height" which is the "standard score"
    calculation of the Height.

2.  Create histograms of both height and standard_height, Make both
    histograms green and with: breaks = 9.

3.  Create a scatter plot of the variables.

## Answer 2

```{r}
#Create the column "Standard_Height" which is the "standard score" calculation of Height. 
info$Standard_Height <- scale(info$Height)
#Another way is with the formula of Standardizing:
info$Standard_Height <- (info$Height - mean(info$Height)) / sd(info$Height)
#Create a histograms of both height and standard_height
hist(info$Height, breaks = 9, col = "seagreen4",
xlab = "Height", main = "Football Players Height Distribution")
hist(info$Standard_Height, breaks = 9, col = "seagreen4",
xlab = "Standardrized Height", main = "Football Players Height Distribution - Standardrized")
#Create a scatter plot of the variables
plot(info$Standard_Height, info$Height,
         main = "Standardrized Height - Height Scatter Plot", xlab = "Standardrized Height", ylab = " Height",
         col = "lightslategrey", pch = 16, cex = 1.5)

```

## Question 3

### 3. As part of our research, we want to check if there are differences between the players from different countries.

1.  By using the ANOVA test, check whether there is a significant
    difference in the height of the players from the different
    countries.

2.  By using the ANOVA test, check whether there is a significant
    difference in the weight of the players from the different
    countries.

3.  By using the ANOVA test, check whether there is a significant
    difference in the age of the players from the different countries.

4.  By using the ANOVA test, check whether there is a significant
    difference in the score of the players from the different countries.

## Answer 3

```{r}
#By using the ANOVA test, check whether there is a significant difference in the height of the players from the different countries.
aov.model = aov(info$Height ~ info$Country)
summary(aov.model)
#The P value is above 0.05 and is therefore not significant

```

```{r}
#By using the ANOVA test, check whether there is a significant difference in the weight of the players from the different countries.
aov.model2 = aov(info$Weight ~ info$Country)
summary(aov.model2)
#The P value is above 0.05 and is therefore not significant

```

```{r}
#By using the ANOVA test, check whether there is a significant difference in the age of the players from the different countries.
aov.model3= aov(info$Age ~ info$Country)
summary(aov.model3)
#The P value is above 0.05 and is therefore not significant

```

```{r}
#By using the ANOVA test, check whether there is a significant difference in the score of the players from the different countries.
aov.model4 = aov(info$Player_score ~ info$Country)
summary(aov.model4)
#The P value is less than 0.05, so it is significant!
#We can use Tukey HSD to check where there is a significant difference!
TukeyHSD(aov.model4)
#There is a significant difference between the Italian soccer players and the English players in terms of scores!

```

## Question 4

### 4. follow the instructions below.

One-sample t-test is used to compare the mean of one sample to a known
standard (or theoretical/hypothetical) mean (μ).

There are 14 Israeli football players in the info dataset with dual
citizenship.

Their weights are: 62, 82, 64, 59, 71, 68, 78, 61, 67, 76, 80, 68, 60,
and 71.

We want to compare their weights to the average of all the players.

1.  Run an F-test for the variables, and compare the Israeli players
    average weight to the players average weight, with a one-sample
    t-test.

## Answer 4

```{r}
#run an F-test for this variables. 
il = c(62, 82, 64, 59, 71, 68, 78, 61, 67, 76, 80, 68, 60, 71)

var.test(il,info$Weight)

```

The p-value of the F-test is = 0.5425. It's greater than the
significance level alpha = 0.05. In conclusion, there is no significant
difference between the variances of the two sets of data. Therefore, we
can use the classic t-test which assumes equality of the two variances.

------------------------------------------------------------------------

```{r}
#compare the Israeli players average weight to the players average weight, with a one-sample t-test.

res = t.test(il, mu=mean(info$Weight))

res


```

The mean of the Israeli players is 69.07143 and all players mean is
75.71551 The p-value of the test is less than the significance level
alpha = 0.05. We can then reject the null hypothesis and conclude that:
The mean weight of the Israeli players is significantly different from
75.71551 with a p-value of 0.006129.

## Lecture 8 Homework

## Question 1

### 1. Follow the instructions below.

An Independent (or unpaired two-sample) t-test is used to compare the
means of two unrelated groups of samples.

In this case, we want to test whether the average age of left foot
players is significantly different from the average age of right foot
players.

1.  Test whether the average age of left players is significantly
    different from the average age of right foot players

## Answer 1

```{r, echo=FALSE}
#Read football_info.csv into R

info = read.csv("football_info.csv")

#Check and clean the data if required

#With summary(info) we can see that there are no NAs, however, there are some wrong values in a few columns.
#The maximum age is 244, which makes no sense, so we need to delete it.
#We can also notice the impossible maximum value of height (281), the maximum value of weight(909), and the minimum value of weight(3).
#So there are some errors in the data and we have to delete the wrong rows.
info = info[order(info$Age),]
info = info[-c(14999,15000),]
info = info[order(info$Height),]
info = info[-14998,]
info = info[order(info$Weight),]
info = info[-14997,]
info = info[-1,]

#Now our data is fully cleaned :)


```

Test whether the average age of left players is significantly different
from the average age of right foot players.

```{r}
res = t.test(Age ~ Foot, data = info)
res
```

The mean age of the left foot players is 24.42190 and the mean age of
right foot players is 26.86195. The p-value of the test is much less
than the significance level alpha = 0.05. We can then reject the null
hypothesis and conclude that: The mean age of the left foot players is
significantly different from right foot players with a p-value of
2.2e-16.

## Question 2

### 2. follow the instructions below.

Paired Student's t-test is used to compare the means of two related
samples.

That is when you have two values (pair of values) for the same samples.

It is known that young football players tend to grow by 3-5 centimeters
after age 16.

1.  With a paired t-test, check if there is a significant difference
    between the height of football players at age 16 and height in
    adulthood.

```{r}
dat = data.frame(player = c(1:15), 
age16 = c(160, 154, 172, 172, 178, 168, 162, 170, 164, 162, 159, 180, 165, 160, 163),
adulthood = c(162, 155, 172, 176, 179, 169, 162, 175, 168, 167, 160, 180, 166, 166, 164))
```

## Answer 2

With a paired t-test, check if there is a significant difference between
the height of football players at age 16 and height in adulthood.

```{r}
dat = data.frame(player = c(1:15), 
age16 = c(160, 154, 172, 172, 178, 168, 162, 170, 164, 162, 159, 180, 165, 160, 163),
adulthood = c(162, 155, 172, 176, 179, 169, 162, 175, 168, 167, 160, 180, 166, 166, 164))

res2 = t.test(dat$age16, dat$adulthood, paired=TRUE)

res2
```

The p-value of the t-test is 0.001316, which is less than the
significance level alpha = 0.05. We can then reject the null hypothesis
and conclude that: The average height of the players at 16 years old is
significantly different from the average height in adulthood with a
p-value = 0.001316.

## Question 3

### 3. Another excellent way to check if the information is normally distributed is by using the Shapiro test.

1.  Run a Shapiro test on the age of all football players, is the age
    distributed normally?

2.  Isolate only the first 30 Italian players, create a histogram of the
    player_score and perform a `shapiro.test`, is the player_score
    normally distributed?

## Answer 3

```{r, eval=FALSE}
shapiro.test(info$Age)
#Error in shapiro.test(info$Age) : sample size must be between 3 and 5000
```

As we can see from the error, the function has a limit of up to 5000
values, and we have much more than that. We cannot change this limit on
the function.

------------------------------------------------------------------------

```{r}
Italy = info[info$Country == "Italy",]
Italy2 = Italy[1:30,]
hist(Italy2$Player_score)
shapiro.test(Italy2$Player_score)
```

Although the histogram does not appear to be normally distributed, the
Shapiro test shows us that your data is normally distributed!

## Bonus question! That is for those who want to go further beyond!

An effective way to remove outliers from data sets in R is to use
"automatic commands" to clean the information.

```{r}
info2 = read.csv("football_info.csv")
x = info2$Height
x_clean =  x[!x %in% boxplot.stats(x)$out] 
```

1.  Run the following code, and by looking at the box-plot of x what
    could be the downside of using "auto clean"?

------------------------------------------------------------------------

```{r, eval=FALSE}

var.test(info$Standard_Height, info$Foot)

```

1.  What can be the problem?

2.  Think of a solution to the problem :)

## Bonus Answer

```{r}
info2 = read.csv("football_info.csv")
x = info2$Height
x_clean =  x[!x %in% boxplot.stats(x)$out] 
par(mfrow=c(1,2))
boxplot(x)
boxplot(x_clean)

```

Despite the great convenience of the automatic cleaning, we have lost
many players whose height is in the area of 160 and above 199, because
of this, it is important that we go through the data first independently
and use automatic tools wisely.

------------------------------------------------------------------------

The Foot column is a charcter. In `var.test(x,y)`; x,y - must to be
numeric vectors To fix this problem you can create a binary column (0/1)
based on a condition in the Foot column.

```{r}
info$Foot[info$Foot == "Left"] = 0  
info$Foot[info$Foot == "Right"] = 1
info$Foot = as.numeric(info$Foot)
#The original column contained text so we needed to convert the column to numeric values
```

Or you can try this:

```{r}
info$Foot_binary = ifelse(info$Foot == "Right", 1, 0)
```

Now lets try the `var.test(x,y)` again!

```{r, eval=FALSE}
var.test(info$Standard_Height, info$Foot)

```
