---
title: "Lecture 9 Exercises"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

This week we will research data from the AirQualityUCI.csv

The dataset contains 827 instances of hourly averaged responses from an
array of 5 metal oxide chemical sensors embedded in an Air Quality
Chemical Multisensor Device. The device was located on the field in a
significantly polluted area, at road level, within an Italian city.

Attribute Information:

1.  Date (DD/MM/YYYY)

2.  Time (HH.MM.SS)

3.  Hourly averaged concentration CO in mg/m\^3

4.  Hourly averaged overall Non Metanic HydroCarbons concentration in
    microg/m\^3

5.  Hourly averaged Benzene concentration in microg/m\^3

6.  Hourly averaged NOx concentration in ppb

7.  Hourly averaged NO2 concentration in microg/m\^3

8.  Temperature in C°

9.  Relative Humidity (%)

10. Absolute Humidity

## Question 1

### 1.1 Follow the instructions below.

1.  Read AirQualityUCI.csv into R.

2.  As always the first we need to do is to check our data! use the
    summary `function` to review all variables and look for NAs and
    outliers.

------------------------------------------------------------------------

Your favorite professor hinted to you that some pollutants are
correlated. Plotting two variables against each other is a great tool to
check their relation. (Feel free to add different colors, sizes, titles,
and trend lines, for the trend line you can use
`abline(lm(x~y, data =))`)

### 1.2 Use plot `(x~y, data =)` on the following columns and examine the results:

1.  CO vs. Benzene

2.  NMHC VS. RH

3.  CO vs. NOx

4.  Temp vs. Nox

5.  Temp vs. RH

## Answer 1.1

```{r}
#Read AirQualityUCI.csv into R

AirQualityUCI = read.csv("AirQualityUCI.csv")

#check all variables with summary

summary(AirQualityUCI)
```

It seems that in this case, there are no significant outliers and no
NAs. Therefore we can move on and use the dataset as it is.

## Answer 1.2

```{r}
#Use plot `(x~y)` on the following columns and examine the results
par(mfrow=c(3,2))

plot(CO ~ Benzene, data = AirQualityUCI)
plot(NMHC ~ RH, data = AirQualityUCI, pch = 19, col = "Darkblue")
plot(CO ~ NOx, data = AirQualityUCI, pch = 16, col = "darkseagreen1")  
abline(lm(CO ~ NOx, data= AirQualityUCI), col = "darkslategrey", lwd = 3)
plot(Temp ~ NOx , data = AirQualityUCI, pch = 8, col = "black")  
abline(lm(Temp ~ NOx, data= AirQualityUCI), col = "orangered1", lwd = 3)
title(main ="Temp vs. NOx", font.main = 2)
plot(Temp ~ RH , data = AirQualityUCI, pch = 16, col = "Grey")  
abline(lm(Temp ~ RH, data= AirQualityUCI), col = "royalblue4", lwd = 4, lty = "dashed")
title(main ="Temp vs. RH", font.main = 4)


```

In the first correlation, there is a very high positive correlation. CO
and Benzene increase together. In the second correlation, It seems that
there is no correlation at all. In the third correlation, there is a
very high positive correlation. CO and NOx increase together. In the
fourth correlation, the line describes the positive trend, however, it
is a very low correlation between the selected variables. In the fifth
correlation there is a moderately strong negative correlation, as long
as Temp increases, RH decreases.

## Question 2

Plotting is a great tool, however, for a better and more accurate
result, we will use a correlation analysis test.

We saw earlier that CO-Benzene relation is high and CO-NOx is high as
well.

### 2. Follow the instructions below.

1.  Check if the attribute above are normally distributed.

2.  Perform correlation tests on CO-Benzene and CO-NOx with the right
    method and answer the question - which relation is stronger?

## Answer 2

Check if the attribute above are normally distributed.

```{r}

shapiro.test(AirQualityUCI$CO)
shapiro.test(AirQualityUCI$Benzene)
shapiro.test(AirQualityUCI$NOx)

```

We see that all the attributes have a P value lower than 0.05, which
means that none of the attributes are normally distributed!

------------------------------------------------------------------------

Also graphically we can see that the selected attributes are distributed
in a way that does not exactly correspond to a normal distribution.

```{r}
par(mfrow=c(1,3))
qqnorm(AirQualityUCI$CO)
qqline(AirQualityUCI$CO,col = "red")
qqnorm(AirQualityUCI$Benzene)
qqline(AirQualityUCI$Benzene,col = "red")
qqnorm(AirQualityUCI$NOx)
qqline(AirQualityUCI$NOx,col = "red")
```

```{r}
par(mfrow=c(1,3))
hist(AirQualityUCI$CO)
hist(AirQualityUCI$Benzene)
hist(AirQualityUCI$NOx)

```

------------------------------------------------------------------------

Perform correlation tests on CO-Benzene and CO-NOx with the right method
and answer the question - which relation is stronger?

```{r}
res1 = cor.test(AirQualityUCI$CO, AirQualityUCI$Benzene, method = "kendall")
res1
```

```{r}
res2 = cor.test(AirQualityUCI$CO, AirQualityUCI$NOx, method = "kendall")
res2
```

Because all the attributes are distributed non-normally, we chose the
Kendall method. The first test shows that the correlation coefficient
between CO and Benzene is 0.8785523 and the p-value is lower than
2.2e-16, thus highly significant. The second test shows that the
correlation coefficient between CO and NOx is 0.8396721 and the p-value
is lower than 2.2e-16, thus highly significant. We can now conclude that
both tests are very highly correlated, however, CO vs. Benzene is
slightly higher correlated.

## Question 3

Carbon monoxide (CO) is a gas that is slightly less dense than air and
consists of one carbon atom and one oxygen atom. Nitrogen dioxide (NO2)
is one of several nitrogen oxides and is considered an air pollutant
created mostly by internal combustion engines (for example car engines).
Your mission is to explain NO2 only with CO, meaning that you should run
a simple linear regression model with NO2 being the outcome variable. In
other words, you want to use CO to predict NO2. Before doing so,
remember that the outcome variable should be normally distributed,

### 3. Follow the instructions below.

1.  Check if the outcome variable (NO2) is normaly distributed.

2.  Check for linearity between the attributes.

3.  Check for multicollinearity (Independence of observations).

4.  Run simple linear regression model as you explain NO2 with CO.

5.  Check for Homoscedasticity (homogeneity of variance).

## Answer 3

```{r}
#Check if the outcome variable (NO2) is normaly distributed
shapiro.test(AirQualityUCI$NO2)

#The p-value (0.4872) is clearly higher than 0.05, so we can conclude it is normally distributed and thus we can run a simple linear regression.
```

------------------------------------------------------------------------

```{r}
#Check for linearity between the attributes.

plot(NO2 ~ CO , data = AirQualityUCI, pch = 16, col = "Grey")  
abline(lm(NO2 ~ CO, data= AirQualityUCI), col = "royalblue4", lwd = 4, lty = "dashed")
title(main ="NO2 vs. CO", font.main = 4)


```

We can see a medium-strong linear relationship between the attributes.

------------------------------------------------------------------------

Check for multicollinearity (Independence of observations).

Because we only have one independent variable and one dependent
variable, we don't need to test for any hidden relationships among
variables.

------------------------------------------------------------------------

```{r}
#Run simple linear regression model as you explain NO2 with CO.
fit = lm(NO2 ~ CO, data=AirQualityUCI)
summary(fit)

```

R-Squared is a metric for evaluating the goodness of fit of your model.
In this case, we can say that \~74% of the cause for NO2 is due to CO
(pretty high).

------------------------------------------------------------------------

Check for homoscedasticity.

```{r}
par(mfrow=c(2,2)) 
plot(fit) 

library(car)
ncvTest(fit)
```

In the Residuals vs Fitted plot, we can see some heteroscedasticity, a
sort of asymmetrical "megaphone" shape appearing. In addition to that by
using ncvTest, we can see a really low P value, the low value means a
lack of homoscedasticity.

Homoscedasticity, or homogeneity of variances, is an assumption of equal
or similar variances in different groups being compared. This is an
important assumption of parametric statistical tests because they are
sensitive to any dissimilarities. Uneven variances in samples result in
biased and skewed test results.

## Question 4

This time the goal of the model is to establish the relationship between
"NO2" as a response variable with "CO", "Benzene", "NOx", "Temp" and
"RH" as predictor variables.

### 4. follow the instructions below.

1.  Check for linearity between the attributes.

2.  Check for multicollinearity (Independence of observations).

3.  Run Multiple Linear Regression to check the relationship between No2
    and the other variables.

4.  Check for Homoscedasticity (homogeneity of variance).

## Answer 4

Check for linearity between the attributes & Check for multicollinearity
(Independence of observations)

```{r, warning=FALSE}
library(corrgram)

cordat= (AirQualityUCI[,c(7,3,5,6,8,9)])
corrgram(cordat)
```

By using the corrgram function, we can see that between NO2 and the
other gases there is a weak to strong correlation, this shows the
linearity of the relationship. However, we have multicollinearity
between the different gases in the linear model. In other words, there
is a strong correlation between the gases and not only with NO2. In such
a case you can use the model for prediction but not for measuring the
influence of the different gasses on NO2.

------------------------------------------------------------------------

```{r}
#Run Multiple Linear Regression to check the relationship between No2 and the other variables.

model <- lm(NO2 ~ CO + Benzene + NOx + Temp + RH , data = AirQualityUCI)
summary(model)

```

The Multiple Linear Regression model has an R-squared of 0.79 which says
that \~79% of the cause of a NO2 is due to the different variables.
Except for RH, all predictors have a very low p-value (good
significance).

------------------------------------------------------------------------

```{r}

par(mfrow=c(2,2)) 
plot(model) 
library(car)
ncvTest(model)

```

In the Residuals vs Fitted plot, we can see some heteroscedasticity, a
sort of asymmetrical "megaphone" shape appearing. In addition to that by
using ncvTest, we can see a really low P value, the low value means a
lack of homoscedasticity.

Homoscedasticity, or homogeneity of variances, is an assumption of equal
or similar variances in different groups being compared. This is an
important assumption of parametric statistical tests because they are
sensitive to any dissimilarities. Uneven variances in samples result in
biased and skewed test results.

## Lecture 9 Homework

## Question 1

### 1. Follow the instructions below.

After modeling NO2 with different variables, it is time to visualize
some connections between them.

1.  NO2 vs. CO

2.  NO2 vs. Benzene

3.  NO2 vs. Temp

4.  NO2 vs. RH

5.  Create one PDF file with 4 plots as explained above

All plots should be with the same plotting symbol "15", in color
"bisque3", 0.5 text size inside the plot, and proper title and axes
labels. Add abline to each plot. all ablines should be colored "brown4"
with a line width of 2.

## Answer 1

```{r}

AirQualityUCI = read.csv("AirQualityUCI.csv")

#Create one PDF file with 4 plots as explained above
    pdf("NO2.pdf") #Calling for PDF file
    par(mfcol = c(2,2)) #Changing the plot panel layout to 2 x 2
    plot(NO2 ~ CO , data = AirQualityUCI,pch = 15, col = "bisque3", cex = 0.5,
         xlab = "CO", ylab = "NO2")  
    title(main ="NO2 vs. CO")
    abline(lm(NO2 ~ CO, data= AirQualityUCI), col = "brown4", lwd = 2)
    plot(NO2 ~ Benzene , data = AirQualityUCI,pch = 15, col = "bisque3", cex = 0.5,
         xlab = "Benzene", ylab = "NO2")  
    title(main ="NO2 vs. Benzene")
    abline(lm(NO2 ~ Benzene, data= AirQualityUCI), col = "brown4", lwd = 2)
    plot(NO2 ~ Temp , data = AirQualityUCI,pch = 15, col = "bisque3", cex = 0.5,
         xlab = "Temp", ylab = "NO2")  
    title(main ="NO2 vs. Temp")
    abline(lm(NO2 ~ Temp, data= AirQualityUCI), col = "brown4", lwd = 2)
    plot(NO2 ~ RH , data = AirQualityUCI,pch = 15, col = "bisque3", cex = 0.5,
         xlab = "RH", ylab = "NO2")  
    title(main ="NO2 vs. RH")
    abline(lm(NO2 ~ RH, data= AirQualityUCI), col = "brown4", lwd = 2)
    dev.off() #Changing the layout back to the default


```

## Question 2

### 2.1 Follow the instructions below.

Read the datafile prawnGR.CSV into R.

The data was collected from an experiment to investigate the difference
in the growth rate of the giant tiger prawn fed either an artificial or
natural diet.

1.  Download and read the datafile prawnGR.CSV into R.

2.  Have a quick look at the structure of this dataset.

3.  plot the growth rate versus the diet using an appropriate plot.

4.  How many observations are there in each diet treatment?

### 2.2

You want to compare the difference in growth rate between the two diets
using a two sample t-test. Before you conduct the test, you need to
assess the normality and equal variance assumptions.

1.  test normality assumption of the variables.

2.  test for equal variance of the variables.

### 2.3

Conduct a two sample t-test using the `t.test()` function . Use the
argument `var.equal = TRUE` to perform the t-test assuming equal
variances.

1.  What is the null hypothesis you want to test?

2.  Do you reject or fail to reject the null hypothesis?

3.  What is the value of the t statistic, degrees of freedom and p
    value?

## Answer 2.1

```{r,eval=FALSE}
prawns = read.csv("prawnGR.CSV", stringsAsFactors = TRUE)

# take a look at the data
str(prawns)

# 'data.frame':   60 obs. of  2 variables:
#  $ GRate: num  9.77 10.29 10.05 10.08 9.31 ...
#  $ diet : Factor w/ 2 levels "Artificial","Natural":...

summary(prawns)

#      GRate                diet   
#  Min.   : 7.395   Artificial:30  
#  1st Qu.: 9.550   Natural   :30  
#  Median : 9.943                  
#  Mean   : 9.920                  
#  3rd Qu.:10.344                  
#  Max.   :11.632         

#How many observations are there in each diet treatment?


table(prawns$diet)

# Artificial    Natural 
#         30         30 


# plot the growth rate versus the diet using an appropriate plot

boxplot(GRate ~ diet, data = prawns, xlab = "Diet", ylab = "Growth Rate")
```

## Answer 2.2

```{r, eval=FALSE}
# test normality assumption

# Do not perform test on all data together, i.e.

shapiro.test(prawns$GRate) # this is wrong!!

# Need to test for departures from normality for each group 
# separately. Remember your indexing [ ]

shapiro.test(prawns$GRate[prawns$diet == "Artificial"])

#         Shapiro-Wilk normality test
# 
# data:  prawns$GRate[prawns$diet == "Artificial"] 
# W = 0.9486, p-value = 0.1553

shapiro.test(prawns$GRate[prawns$diet == "Natural"])

#         Shapiro-Wilk normality test
# 
# data:  prawns$GRate[prawns$diet == "Natural"] 
# W = 0.9598, p-value = 0.307

# Therefore no evidence to reject the Null hypothesis and data are normally distributed

# However much better assessment of normality is to use a quantile - quantile plot
# looking for points to lie along the line for normality

qqnorm(prawns$GRate[prawns$diet == "Artificial"])
qqline(prawns$GRate[prawns$diet == "Artificial"])

qqnorm(prawns$GRate[prawns$diet == "Natural"])
qqline(prawns$GRate[prawns$diet == "Natural"])

# test for equal variance
# if normal
# Null hypothesis Ho: variances are equal

var.test(prawns$GRate ~ prawns$diet)

#         F test to compare two variances
# data:  prawns$GRate by prawns$diet 
# F = 1.9629, num df = 29, denom df = 29, p-value = 0.07445
# alternative hypothesis: true ratio of variances is not equal to 1 
# 95 percent confidence interval:
#  0.9342621 4.1240043 
# sample estimates:
# ratio of variances 
#           1.962881 

# No evidence to reject null hypothesis (P=0.07) therefore no 
# difference in variance
```

## Answer 2.3

```{r, eval=FALSE}
# conduct t-test assuming equal variances
# Null hypothesis Ho: no difference in growth rate 
# between prawns fed on artificial diet or Natural diet

t.test(GRate ~ diet, var.equal = TRUE, data = prawn)

#         Two Sample t-test
# 
# data:  prawns$GRate by prawns$diet 
# t = -1.3259, df = 58, p-value = 0.1901
# alternative hypothesis: true difference in means is not equal to 0 
# 95 percent confidence interval:
#  -0.6319362  0.1283495 
# sample estimates:
# mean in group Artificial    mean in group Natural 
#                 9.794133                10.045927 
# 

# No evidence to reject the Null hypothesis, therefore no 
# difference in growth rate of prawns fed on either artificial 
# or natural diet (t = -1.33, df = 58, p = 0.19).
```

## Question 3

### 3. Follow the instructions below.

An alternative (but equivalent) way to compare the mean growth rate
between diets is to use a linear model.

1.  Use the `lm()` function to fit a linear model with GRate as the
    response variable and diet as an explanatory variable.

2.  Produce an ANOVA table for the fitted model using the `anova()`
    function.

3.  Compare the ANOVA p value to the p value obtained using a t-test.
    What do you notice?

## Answer 3

```{r, eval=FALSE}
# fit the model

growth.lm <- lm(GRate ~ diet, data = prawns)
```

```{r, eval=FALSE}
# produce the ANOVA table

anova(growth.lm)

# Analysis of Variance Table
# 
# Response: GRate
#           Df Sum Sq Mean Sq F value Pr(>F)
# diet       1  0.951 0.95100  1.7579 0.1901
# Residuals 58 31.377 0.54098  

# notice the p value is the same as for the t-test
# also if you square the t statistic from the t-test
# you get the F statistic from the linear model.
# They're the same test
```

## Bonus question! That is for those who want to go further beyond!

In 1973, Francis Anscombe famously created four datasets with remarkably
similar numerical properties, but obviously different graphic
relationships. The Anscombe dataset contains the x and y coordinates for
these four datasets, along with a grouping variable, set, that
distinguishes the quartet.

It may be helpful to remind yourself of the graphic relationship by
viewing the four scatterplots:

```{r}

dat <- datasets::anscombe
Anscombe <- data.frame(
    set  = rep(1:4, each = 11),
    x = unlist(dat[ ,c(1:4)]),
    y = unlist(dat[ ,c(5:8)])
    )
rownames(Anscombe) <- NULL

Anscombe1 = Anscombe[Anscombe$set == 1,]
Anscombe2 = Anscombe[Anscombe$set == 2,]
Anscombe3 = Anscombe[Anscombe$set == 3,]
Anscombe4 = Anscombe[Anscombe$set == 4,]

par(mfcol = c(2,2))
plot(Anscombe1$x,Anscombe1$y)
plot(Anscombe2$x,Anscombe2$y)
plot(Anscombe3$x,Anscombe3$y)
plot(Anscombe4$x,Anscombe4$y)


```

For each of the four sets of data points in the Anscombe dataset,
compute the following in the order specified.

1.  Mean of x.

2.  Standard deviation of x.

3.  Mean of y.

4.  Standard deviation of y.

5.  Correlation coefficient between x and y.

## Bonus Answer

```{r, echo=FALSE}

dat <- datasets::anscombe
Anscombe <- data.frame(
    set  = rep(1:4, each = 11),
    x = unlist(dat[ ,c(1:4)]),
    y = unlist(dat[ ,c(5:8)])
    )
rownames(Anscombe) <- NULL

Anscombe1 = Anscombe[Anscombe$set == 1,]
Anscombe2 = Anscombe[Anscombe$set == 2,]
Anscombe3 = Anscombe[Anscombe$set == 3,]
Anscombe4 = Anscombe[Anscombe$set == 4,]

par(mfcol = c(2,2))
plot(Anscombe1$x,Anscombe1$y)
plot(Anscombe2$x,Anscombe2$y)
plot(Anscombe3$x,Anscombe3$y)
plot(Anscombe4$x,Anscombe4$y)


```

```{r}
#Mean of x:

mean(Anscombe1$x)
mean(Anscombe2$x)
mean(Anscombe3$x)
mean(Anscombe4$x)

```

------------------------------------------------------------------------

```{r}
#Mean of y:

mean(Anscombe1$y)
mean(Anscombe2$y)
mean(Anscombe3$y)
mean(Anscombe4$y)

```

------------------------------------------------------------------------

```{r}
#Standard deviation of x.
sd(Anscombe1$x)
sd(Anscombe2$x)
sd(Anscombe3$x)
sd(Anscombe4$x)
```

------------------------------------------------------------------------

```{r}
#Standard deviation of y.
sd(Anscombe1$y)
sd(Anscombe2$y)
sd(Anscombe3$y)
sd(Anscombe4$y)
```

------------------------------------------------------------------------

```{r}
#Correlation coefficient between x and y
cor(Anscombe1$x,Anscombe1$y)
cor(Anscombe2$x,Anscombe2$y)
cor(Anscombe3$x,Anscombe3$y)
cor(Anscombe4$x,Anscombe4$y)

```

------------------------------------------------------------------------

```{r}
#Correlation coefficient between x and y
cor(Anscombe1$x,Anscombe1$y, method = "kendall")
cor(Anscombe2$x,Anscombe2$y,  method = "kendall")
cor(Anscombe3$x,Anscombe3$y,  method = "kendall")
cor(Anscombe4$x,Anscombe4$y,  method = "kendall")

```

------------------------------------------------------------------------

```{r}
#Correlation coefficient between x and y
cor(Anscombe1$x,Anscombe1$y, method = "spearman")
cor(Anscombe2$x,Anscombe2$y,  method = "spearman")
cor(Anscombe3$x,Anscombe3$y,  method = "spearman")
cor(Anscombe4$x,Anscombe4$y,  method = "spearman")

```

All the plots had the same mean and standard deviation, and even in the
standard correlation test, all graphs had the same value. That is why it
is important to remember to use the Kendall and Spearman method when the
data is not normally distributed.
