---
title: "Lecture 8 Homework"
format:
  html: 
    theme: literal
---

## Question 1

An Independent (or unpaired two-sample) t-test is used to compare the means of two unrelated groups of samples.

In this case, we want to test whether the average age of left foot players is significantly different from the average age of right foot players.

-   Test whether the average age of left players is significantly different from the average age of right foot players

## Question 2

Paired Student's t-test is used to compare the means of two related samples.

That is when you have two values (pair of values) for the same samples.

It is known that young football players tend to grow by 3-5 centimeters after age 16.

-   With a paired t-test, check if there is a significant difference between the height of football players at age 16 and height in adulthood.

```{r}
dat = data.frame(player = c(1:15), 
age16 = c(160, 154, 172, 172, 178, 168, 162, 170, 164, 162, 159, 180, 165, 160, 163),
adulthood = c(162, 155, 172, 176, 179, 169, 162, 175, 168, 167, 160, 180, 166, 166, 164))
```

## Question 3

Another excellent way to check if the information is normally distributed is by using the Shapiro test.

-   Run a Shapiro test on the age of all football players, is the age distributed normally?

-   Isolate only the first 30 Italian players, create a histogram of the player_score and perform a `shapiro.test`, is the player_score normally distributed?

## Bonus question! That is for those who want to go further beyond!

An effective way to remove outliers from data sets in R is to use "automatic commands" to clean the information.

```{r}
info2 = read.csv("football_info.csv")
x = info2$Height
x_clean =  x[!x %in% boxplot.stats(x)$out] 
```

-   Run the following code, and by looking at the box-plot of x what could be the downside of using "auto clean"?

------------------------------------------------------------------------

```{r, eval=FALSE}

var.test(info$Standard_Height, info$Foot)

```

-   What can be the problem?

-   Think of a solution to the problem :)
