---
title: "Lecture 8 Homework Answers"
format:
  html: 
    theme: literal
---

## Answer 1

```{r, echo=FALSE}
#Read football_info.csv into R

info = read.csv("football_info.csv")

#Check and clean the data if required

#With summary(info) we can see that there are no NAs, however, there are some wrong values in a few columns.
#The maximum age is 244, which makes no sense, so we need to delete it.
#We can also notice the impossible maximum value of height (281), the maximum value of weight(909), and the minimum value of weight(3).
#So there are some errors in the data and we have to delete the wrong rows.
info = info[order(info$Age),]
info = info[-c(14999,15000),]
info = info[order(info$Height),]
info = info[-14998,]
info = info[order(info$Weight),]
info = info[-14997,]
info = info[-1,]

#Now our data is fully cleaned :)


```


Test whether the average age of left players is significantly different from the average age of right foot players.

```{r}
res = t.test(Age ~ Foot, data = info)
res
```

The mean age of the left foot players is 24.42190 and the mean age of right foot players is 26.86195. The p-value of the test is much less than the significance level alpha = 0.05. We can then reject the null hypothesis and conclude that: The mean age of the left foot players is significantly different from right foot players with a p-value of 2.2e-16.
   
    
  
## Answer 2

With a paired t-test, check if there is a significant difference between the height of football players at age 16 and height in adulthood.

```{r}
dat = data.frame(player = c(1:15), 
age16 = c(160, 154, 172, 172, 178, 168, 162, 170, 164, 162, 159, 180, 165, 160, 163),
adulthood = c(162, 155, 172, 176, 179, 169, 162, 175, 168, 167, 160, 180, 166, 166, 164))

res2 = t.test(dat$age16, dat$adulthood, paired=TRUE)

res2
```

The p-value of the t-test is 0.001316, which is less than the significance level alpha = 0.05.
We can then reject the null hypothesis and conclude that:
The average height of the players at 16 years old is significantly different from the average height in adulthood with a p-value = 0.001316.


## Answer 3

```{r, eval=FALSE}
shapiro.test(info$Age)
#Error in shapiro.test(info$Age) : sample size must be between 3 and 5000
```

As we can see from the error, the function has a limit of up to 5000 values, and we have much more than that. We cannot change this limit on the function.

___

```{r}
Italy = info[info$Country == "Italy",]
Italy = Italy[1:30,]
hist(Italy$Player_score)
shapiro.test(Italy$Player_score)
```

Although the histogram does not appear to be normally distributed, the Shapiro test shows us that your data is normally distributed!


## Bonus Answer

```{r}
info2 = read.csv("football_info.csv")
x = info2$Height
x_clean =  x[!x %in% boxplot.stats(x)$out] 
par(mfrow=c(1,2))
boxplot(x)
boxplot(x_clean)

```

Despite the great convenience of the automatic cleaning, we have lost many players whose height is in the area of 160 and above 199, because of this, it is important that we go through the data first independently and use automatic tools wisely.

___

The Foot column is a charcter. In `var.test(x,y)`; x,y - must to be numeric vectors
To fix this problem you can create a  binary column (0/1) based on a condition in the Foot column.

```{r}
info$Foot[info$Foot == "Left"] = 0  
info$Foot[info$Foot == "Right"] = 1
info$Foot = as.numeric(info$Foot)
#The original column contained text so we needed to convert the column to numeric values
```

Or you can try this:

```{r}
info$Foot_binary = ifelse(info$Foot == "Right", 1, 0)
```

Now lets try the `var.test(x,y)` again!

```{r, eval=FALSE}
var.test(info$Standard_Height, info$Foot)

```