---
title: "Lecture 10 Homework"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

## Question 1:

An SF layer is a special type of data.frame, where one of the columns is
a geometry column. Therefore, many of the functions we learned when
working with data frames also work on sf layers.

Run the following code:

```{r, message=FALSE,warning=FALSE}

require(sf)

rainfall = read.csv("rainfall.csv")

rainfall = st_as_sf(rainfall, coords = c("x_utm", "y_utm"), crs = 32636)

plot(st_geometry(rainfall))

```

-   Plot the stations that had more than 150 mm of rain in January.

-   Plot the stations that had less than 5 mm of rain in April.

-   Find the station that had the lowest mm of rain in December, and
    print the station name.

-   Find the station that had the highest mm of rain in February, and
    print the station name.

## Question 2:

-   Check the rain averages in the winter months - December, January,
    and February, is one of the winter months significantly different in
    the amount of rainfall?

-   Check the rainfall averages in the spring months - March, April, and
May, is one of the summer months significantly different in the
amount of rain?

## Bonus question! That is for those who want to go further beyond!

Create the value and time vectors representing the water level of the
BGU river in January and August during 2000-2006.

```{r}
value = c(1,0.8,1.2,0.5,1.8,1.5,1.1,0.8,1.3,0.3,1,0.8,1.5,1.1,1.9,1.2)
time = rep(2000:2006, each = 2)
time = rep(2000:2006, each = 2)
time = paste0(time, c("-01-30", "-08-30"))
time = as.Date(time)
```

-   Plot the water level in January (month "01") as a function of the
    water level in August (month "08"), between 2000 and 2006, where
    each point represents one year. The points need to be connected in
    chronological order.

-   Add text labels marking the water levels in 2000 and 2006.

You should get something like that:

```{r,echo=FALSE}
value = c(1,0.8,1.2,0.5,1.8,1.5,1.1,0.8,1.3,0.3,1,0.8,1.5,1.1,1.9,1.2)
time = rep(2000:2006, each = 2)
time = rep(2000:2006, each = 2)
time = paste0(time, c("-01-30", "-08-30"))
time = as.Date(time)

month = as.numeric(as.character(time, "%m"))

year = as.numeric(as.character(time, "%Y"))

Jan = value[month == 1]

Aug = value[month == 8]

plot(Jan, Aug, xlab = "January", ylab = "August", type = "b")
text(Jan[1], Aug[1], labels = year[which.min(year)], pos = 3, col = "red")
text(Jan[length(Jan)], Aug[length(Aug)],labels = year[which.max(year)], pos = 3, col = "red")



```
