---
title: "Lecture 10 Homework Answers"
format:
  html: 
    theme: literal
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}

require(sf)

rainfall = read.csv("rainfall.csv")

rainfall = st_as_sf(rainfall, coords = c("x_utm", "y_utm"), crs = 32636)


```


## Answer 1

Plot the stations that had more than 150 mm of rain in January.

```{r}
Jan_rain = rainfall[rainfall$jan > 150,]
plot(st_geometry(Jan_rain))

```


Plot the stations that had less than 5 mm of rain in April.

```{r}
April_rain = rainfall[rainfall$apr < 5,]
plot(st_geometry(April_rain))

```

Find the station that had the lowest mm of rain in December, and print the station name.

```{r}
Dec_rain = rainfall[which.min(rainfall$dec),]
print(Dec_rain$name)

```

Find the station that had the highest mm of rain in February, and print the station name.


```{r}
Feb_rain = rainfall[which.max(rainfall$feb),]
print(Feb_rain$name)

```


## Answer 2

Check the rain averages in the winter months - December, January,
and February, is one of the winter months significantly different in
the amount of rainfall?


```{r}

Dec = rainfall$dec
Jan = rainfall$jan
Feb = rainfall$feb

# organize the data and the grouping variable as a data frame
winter_rain = data.frame(
  rain = c(Dec, Jan, Feb),
  group = rep(c("December", "January", "February"), each=169))

## fit linear model and perform anova
aov.model = aov(winter_rain$rain ~ winter_rain$group)
summary(aov.model)

TukeyHSD(aov.model)


#January-December & January-February are significantly different int the amount of rainfall.

  
```

___

Check the rainfall averages in the spring months - March, April, and
May, is one of the summer months significantly different in the
amount of rain?

```{r}

March = rainfall$mar
April = rainfall$apr
May = rainfall$may

# organize the data and the grouping variable as a data frame
spring_rain = data.frame(
  rain = c(March, April, May),
  group = rep(c("March", "April", "May"), each=169))

## fit linear model and perform anova
aov.model2 = aov(spring_rain$rain ~ spring_rain$group)
summary(aov.model2)

TukeyHSD(aov.model2)


# All the months are significantly different in the amount of rainfall.

  
```


## Bonus Answer
```{r}
value = c(1,0.8,1.2,0.5,1.8,1.5,1.1,0.8,1.3,0.3,1,0.8,1.5,1.1,1.9,1.2)
time = rep(2000:2006, each = 2)
time = rep(2000:2006, each = 2)
time = paste0(time, c("-01-30", "-08-30"))
time = as.Date(time)

month = as.numeric(as.character(time, "%m"))

year = as.numeric(as.character(time, "%Y"))

Jan = value[month == 1]

Aug = value[month == 8]

plot(Jan, Aug, xlab = "January", ylab = "August", type = "b")
text(Jan[1], Aug[1], labels = year[which.min(year)], pos = 3, col = "red")
text(Jan[length(Jan)], Aug[length(Aug)],labels = year[which.max(year)], pos = 3, col = "red")



```

